---
output:
  pdf_document:
    citation_package: natbib
    latex_engine: pdflatex
    toc: false
    keep_tex: true
title: " Robust Empirical Bayes Confidence Intervals"
author: "Michal Kolesár"
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: mathpazo
bibliography: library.bib
fontsize: 11pt
vignette: >
  %\VignetteIndexEntry{ShiftShareSE}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE, cache=FALSE}
library("knitr")
knitr::opts_knit$set(self.contained = FALSE)
knitr::opts_chunk$set(tidy = TRUE, collapse=TRUE, comment = "#>",
                      tidy.opts=list(blank=FALSE, width.cutoff=55))
```

The package `ebci` implements robust empirical Bayes confidence intervals
(EBCIs) proposed by @akm20 for inference in a normal means model $Y_i\sim
N(\theta_i, \sigma^2_i)$, $i=1, \dotsc, n$.

# Setup

Suppose we use an empirical Bayes estimator of $\theta_i$ that shrinks toward
the predictor based on the regression of $\theta_i$ onto $X_i$ (equivalently,
regression of $Y_i$ onto $X_i$),
\begin{equation}\label{Bayes-estimator}
\hat{\theta}_{i} = X_i'\delta+w_i(Y_i-X_i'\delta),
\end{equation}
where $\delta=E[X_i X_i']^{-1}E[X_i\theta_i]$,
$w_i=\frac{\mu_{2}}{\mu_{2}+\sigma_{i}^{2}}$, and
\begin{equation}\label{mu2-independence}
\mu_{2}=E[(\theta_i-X_i'\delta)^{2} \mid X_{i}, \sigma_i].
\end{equation}
We assume that $\mu_2$ doesn't depend on $\sigma_i$. @morris83 proposes to use the
*parametric EBCI*
\begin{equation*}
\hat{\theta}_{i} \pm \frac{z_{1-\alpha/2}}{\sqrt{w_i}}w_i\sigma_i.
\end{equation*}

The critical value $z_{1-\alpha/2}/\sqrt{w_i}$ is larger than the usual critical
value $z_{1-\alpha/2}=$`qnorm(1-alpha/2)` if the estimator was unbiased conditional on
$\theta_i$. This CI is justified if we strengthen the assumption
(\ref{mu2-independence}) by making the normality assumption $\theta_i\mid X_{i},
\sigma_{i} \sim N(X_i'\delta, \mu_2)$: to account for the bias
$b_i=(1-w)(\theta_i-X_i'\delta)$ of the estimator (more precisely $b_i$ is the bias conditional on $(X_i,\theta_i,\sigma_i)$), the parametric EBCI assumes that
it is normally distributed.

A *robust EBCI* that is only uses (\ref{mu2-independence}) and not the normality
assumption takes the form
\begin{equation}\label{robust-ebci}
X_i'\delta+w_i(Y_i-X_i'\delta) \pm cva_{\alpha}(m_2, \infty) w_i\sigma_i,
\,\, m_2=(1-1/w_i)^2\mu_2/\sigma^2_i,
\end{equation}
where the critical value $cva_{\alpha}$ is derived in @akm20. Here $m_2$ is the
second moment of the bias-standard deviation ratio $m_2=b_i/(w_i\sigma_i)$,
which we refer to as the normalized bias. This critical value imposes a
constraint (\ref{mu2-independence}) on the second moment of $\theta_i$, but no
constraints on higher moments. We can make the critical value smaller by also
imposing a constraint on the kurtosis of $\theta_i$ (or equivalently, the
kurtosis of the normalized bias)
\begin{equation}
\label{kappa-independence}
 \kappa=E[(\theta_i-X_i'\delta)^{4}
\mid X_{i}, \sigma_i]/\mu_{2}^2=E[b_i^4]/E[b_i^2]^2.
\end{equation}
In analogy to (\ref{mu2-independence}), we assume here that the conditional fourth
moment of $\theta_i-X_i'\delta$ doesn't depend on $(X_i,\sigma_i)$. In this
case, the robust EBCI takes the form
\begin{equation*}
\hat{\theta}_{i} \pm cva_{\alpha}(m_2,\kappa)w_i\sigma_i,\,\, m_2=(1-1/w_i)^2\mu_2\sigma^2_i,
\end{equation*}

These critical values are implemented in the package by the `cva` function:

```{r}
library("ebci")
## If m_2=0, then we get the usual critical value
cva(m2=0, kappa=Inf, alpha=0.05)$cv
## Otherwise the critical value is larger:
cva(m2=4, kappa=Inf, alpha=0.05)$cv
## Imposing a constraint on kurtosis tightens it
cva(m2=4, kappa=3, alpha=0.05)$cv
```

# Example

Here we illustrate the use of the package using a dataset from @ChHe18ii (CH
hereafter). The dataset is included in the package as the list `cz`. Run
`?cz` for a full description of the dataset.

```{r}
## For illustration, only use 20 largest CZ
df <- cz[sort(cz$pop, index.return=TRUE, decreasing=TRUE)$ix[1:20], ]

## As Y_i, use fixed effect estimate theta25 of causal effect of neighborhood for children with parents at the 25th percentile of income distribution. The standard error for this estimate is se25. As predictors use average outcome for permanent residents (stayers), stayer25. Let us use 90% CIs, and let us set kappa=Inf (so we only don't use a constraint on kurtosos) in order to decrease the computational time. Given this, we do not need to parallelize the computation, and can set cores=1:
r <- ebci(formula=theta25~stayer25, data=df, se=se25, alpha=0.1, cores=1, kappa=Inf)
```

For shrinkage toward the grand mean, or toward zero, use the specification
`theta25 ~ 1`, or `theta25 ~ 0` in the `formula` argument of `ebci`

The return value contains:

1. The least squares estimate of $\delta$:
   ```{r}
   r$delta
   ```
2. Estimates of $\sqrt{\mu_2}$ and $\kappa$, obtained using method of moments. We
   set `kappa=Inf`, so the kurtosis value is taken to be infinity:
   ```{r}
   c(r$sqrt_mu2, r$kappa)
   ```
3. A data frame with columns:
   ```{r}
   names(r$df)
   ```

The columns of the data frame refer to:

- `w_eb` Empirical Bayes shrinkage factor $w_i=\mu_2/(\mu_2+\sigma_i^2)$.
- `th_eb` Empirical Bayes estimator $\hat{\theta_i}$ given in (\ref{Bayes-estimator})
- `len_eb` Half-length $cva_{\alpha}(m_2, \kappa)w_i\sigma_i$ of the robust
  EBCI, so that the lower endpoint of the EBCIs are given by `th_eb-len_eb`, and
  the upper endpoint by `th_eb+len_eb`. For a given observation, this can be
  also computed directly using the `cva` function:
  ```{r}
  cva(m2=((1-1/r$df$w_eb[1])*r$sqrt_mu2/r$df$se[1])^2, Inf, alpha=0.1)$cv*
  r$df$w_eb[1]*r$df$se[1]
  r$df$len_eb[1]
  ```
- `len_pa` Half-length $z_{1-\alpha/2}\sqrt{w_i}\sigma_i$ of the parametric EBCI.
- `w_opt` Shrinkage factor that optimizes the length of the resulting confidence
  interval, that is, the value of $w_i$ that minimizes (\ref{robust-ebci}) over
  $w_i$
- `th_op` Estimator based on the length-optimal shrinkage factor `w_opt`
- `len_op` Half-length $cva_{\alpha}(m_2, \kappa)w_i\sigma_i$ of the
  length-optimal EBCI.
- `th_us` The unshrunk estimate $Y_i$, as specified in the `formula` argument of
  the function `ebci`.
- `len_us` Half-length $z_{1-\alpha/2}\sigma_i$ of the CI based on the unshrunk
  estimate
- `se` The standard error $\sigma_i$, as specified by the argument `se` of the
  `ebci` function.
- `ncov_pa` Maximal non-coverage of the parametric EBCI.

Using the data frame, we can give a table summarizing the results:
```{r}
knitr::kable(data.frame(name=paste0(df$czname, ", ", df$state), estimate=r$df$th_eb,
lower_ci=r$df$th_eb-r$df$len_eb, upper_ci=r$df$th_eb+r$df$len_eb), digits=3)
```

An analogous table, but constructed using length-optimal shrinkage:
```{r}
knitr::kable(data.frame(name=paste0(df$czname, ", ", df$state), estimate=r$df$th_op,
lower_ci=r$df$th_op-r$df$len_op, upper_ci=r$df$th_op+r$df$len_op), digits=3)
```

Using length-optimal shrinkage tightens the robust EBCIs by about 6 percentage points on average
```{r}
mean(r$df$len_op)/mean(r$df$len_eb)
```
On the other hand, using the parametric EBCI yields CIs that mildly violate the
90% coverage requirement:
```{r}
mean(r$df$ncov_pa)
```

# References
